\textbf{Papers list (so far, not everything)}
\begin{itemize}
    \item 2017 Image-to-Image Translation with Conditional Adversarial Networks (Pix2Pix)
    \url{https://arxiv.org/abs/1611.07004}

    \item 2017 Unpaired Image-to-Image Translation using Cycle-Consistent Adversarial Networks (CycleGAN)
    \url{https://arxiv.org/abs/1703.10593}

    \item 2018 THE SEN1-2 DATASET FOR DEEP LEARNING IN SAR-OPTICAL DATA FUSION
    \url{https://isprs-annals.copernicus.org/articles/IV-1/141/2018/}

    \item 2018 A Conditional Generative Adversarial Network to Fuse Sar And Multispectral Optical Data For Cloud Removal From Sentinel-2 Images
    \url{https://ieeexplore.ieee.org/document/8519215}

    \item 2019 SEN12MS – A CURATED DATASET OF GEOREFERENCED MULTI-SPECTRAL SENTINEL-1/2 IMAGERY FOR DEEP LEARNING AND DATA FUSION
    \url{https://isprs-annals.copernicus.org/articles/IV-2-W7/153/2019/}

    \item 2019 Synthesis of Multispectral Optical Images From SAR/Optical Multitemporal Data Using Conditional Generative Adversarial Networks
    \url{https://ieeexplore.ieee.org/document/8637007}

    \item 2020 Cloud removal in Sentinel-2 imagery using a deep residual neural network and SAR-optical data fusion
    \url{https://www.sciencedirect.com/science/article/pii/S0924271620301398}

    \item 2020 GAN generation of synthetic multispectral satellite images
    \url{https://www.researchgate.net/publication/346821973_GAN_generation_of_synthetic_multispectral_satellite_images}


    
    \item 2020 Cloud Removal for Remote Sensing Imagery via Spatial Attention Generative Adversarial Network
    \url{https://arxiv.org/abs/2009.13015}

    \item 2020 Denoising Diffusion Probabilistic Models (DDPM)
    \url{https://arxiv.org/abs/2006.11239}

    \item 2020 An Image is Worth 16x16 Words: Transformers for Image Recognition at Scale (ViT)
    \url{https://arxiv.org/abs/2010.11929}

    \item 2022 SAR-to-optical image translation by a variational generative adversarial network
    \url{https://www.tandfonline.com/doi/full/10.1080/2150704X.2022.2068986}

    \item 2022 Cloud Removal in Remote Sensing Images Using Generative Adversarial Networks and SAR-to-Optical Image Translation
    \url{https://ieeexplore.ieee.org/document/9627647}

    \item 2023 UnCRtainTS: Uncertainty Quantification for Cloud Removal in Optical Satellite Time Series
    \url{https://arxiv.org/abs/2304.05464}
    
    \item 2023 SAR-to-Optical Image Translation and Cloud Removal Based on Conditional Generative Adversarial Networks: Literature Survey, Taxonomy, Evaluation Indicators, Limits and Future Directions
    \url{https://www.mdpi.com/2072-4292/15/4/1137}

    \item 2023 Mamba: Linear-Time Sequence Modeling with Selective State Spaces
    \url{https://arxiv.org/abs/2312.00752}
------------------------------------
    
    \item 2024 Assessing the Potential of Multi-Temporal Conditional Generative Adversarial Networks in SAR-to-Optical Image Translation for Early-Stage Crop Monitoring
        \url{https://www.mdpi.com/2072-4292/16/7/1199}
    

    \item 2024 DiffCR: A Fast Conditional Diffusion Framework for Cloud Removal from Optical Satellite Images
    \url{https://arxiv.org/abs/2308.04417}

    \item 2024  Conditional Diffusion for SAR to Optical Image Translation
    \url{https://ieeexplore.ieee.org/document/10330015}



    \item 2024  SAR to Optical Image Translation with Color Supervised Diffusion Model
    \url{https://arxiv.org/abs/2407.16921}
    

    \item 2024  High-Resolution SAR-to-Multispectral Image Translation Based on S2MS-GAN
    \url{https://www.mdpi.com/2072-4292/16/21/4045}

    
    \item 2025 SAR-to-Optical Image Translation Using Vision Transformer-Based CGAN
    \url{https://ieeexplore.ieee.org/document/10948876}
    

    \item Huang, B., Li, Y., Han, X. Y., Cui, Y., Li, W., & Li, R. (2015). Cloud removal from optical satellite imagery with SAR imagery using sparse representation. \textit{IEEE Geoscience and Remote Sensing Letters}, 12, 1046–1050.

    \item Xu, M., Jia, X., Pickering, M., & Plaza, A. J. (2016). Cloud removal based on sparse representation via multitemporal dictionary learning. \textit{IEEE Transactions on Geoscience and Remote Sensing}, 54(5), 2998–3006. doi:10.1109/TGRS.2015.2509860

    \item Goodfellow, I., et al. (2014). Generative Adversarial Networks. [*Include full venue details*]

    \item Mirza, M., & Osindero, S. (2014). Conditional Generative Adversarial Nets. [*Include full venue details*]

    \item Enomoto, R., et al. (2017). Cloud removal from RGB satellite images by multispectral conditional GAN using NIR as auxiliary input. [*Complete citation details—journal name*, *volume*, *pages*]

    \item Fuentes Reyes, M., Auer, S., Merkle, N., Henry, C., & Schmitt, M. (2019). SAR-to-Optical Image Translation Based on Conditional Generative Adversarial Networks—Optimization, Opportunities and Limits. \textit{Remote Sensing}, 11. doi:10.3390/rs11172067

    \item Wang, L., Xu, X., Yu, Y., Yang, R., Gui, R., Xu, Z., & Pu, F. (2019). SAR-to-optical image translation using supervised cycle-consistent adversarial networks. \textit{IEEE Access}, 7, 129136–129149. doi:10.1109/ACCESS.2019.2939649

    \item Gao, [Authors], et al. (2020). Cloud removal with fusion of high-resolution optical and SAR images using GAN. \textit{Remote Sensing}, 12(1), 191. doi:10.3390/rs12010191

    \item Naderi Darbaghshahi, F., & Mohammadi, M. R. (2021). Cloud removal via Two-GAN with dilated residual inception blocks (DRIBs). [*Provide journal or conference name, volume, pages *]

    
\end{itemize}
